\section*{Acknowledgements}

We would like to thank the SAMueL project team (Lauren Asare, Julia Frost, Iain Lang, Kristin Liabo, Peter McMeekin, Keira Pratt-Boyden, Cathy Pope, Ken Stein, Penny Thompson, Rachel Jarvie) for their input into this work. We would also like to thank our Patient and Carer Involvement team led by Leon Farmer (David Burgess, Simon Douglas, Ian Hancock, Nicola Hancock, John Williams), and our expert advisory group (Ajay Bhalla, Gary Ford, Martin Utley).

\section*{Declaration of conflicting interests}

The authors declare no potential conflicts of interest with respect to the research, authorship, and/or publication of this article. 

\section*{Funding}

The authors disclose receipt of the following financial support for the research, authorship, and/or publication of this article: This research was funded by the National Institute for Health Research Applied Research Collaboration South West Peninsula and by the National Institute for Health Research Health and Social Care Delivery Research (HSDR) Programme [NIHR134326]. The views expressed in this publication are those of the authors and not necessarily those of the National Institute for Health Research or the Department of Health and Social Care. 

\section*{Informed consent and Ethics}

As we are using anonymised secondary data, collected for national audit, individual consent is not required. SSNAP has approval under section 251 of the NHS Health and Social Care Act (2006) to collect patient level data on the first six months of patient care (ECC 6- 02(FT3)/2012), without requiring individual patient consent. Access to SSNAP data is managed by the UK Healthcare Quality Improvement Partnership (HQIP), with this project being approved by HQIP (HQIP303). More information on SSNAPs use of patient data may be found at: \url{https://www.strokeaudit.org/ SupportFiles/Documents/Patient-area-documents/Fair-processingstatement-for-patients-v7-0.aspx}

As we are using anonymised secondary data, collected for national audit, used for service evaluation and improvement, no ethical approval is required (confirmed using the NHS Health Research Authority decision aid: https://www.hra-decisiontools.org.uk/ethics/).

\section*{Guarantor}

The corresponding author, Michael Allen, is the guarantor of the paper (PI on the NIHR project funding).

